---
title: "FRAUD DETECTION"
output: word_document
---

DATASET USED: 

#IEEE-CIS Fraud Detection Dataset.

The dataset used in this analysis comes from Kaggle competition, IEEE Fraud Detection-CIS. The data on kaggle is broken into two files identity and transaction, which are joined by TransactionID. 

The purpose of this analysis is to predict the probability that an online transaction is fraudulent, as denoted by the binary target isFraud.

The following is a summary of features of the dataset.

Outcome variable :
  
  - isFraud : Binary target variable (0 : Legitimate, 1 : Fraud)

Categorical Features : 
  
  -  ProductCD: product code, the product for each transaction.
-  card1 - card6: payment card information, such as card type, card category, issue bank,             country, etc.
-  addr1, addr2: address
-  P_emaildomain: purchaser email domain
-  R_emaildomain: recipient email domain
-  M1-M9: match, such as names on card and address, etc.
-  DeviceType
-  DeviceInfo
-  id_12 - id_38 : identity information – network connection information (IP, ISP, Proxy, etc)        and digital signature (UA/browser/os/version, etc) associated with transactions. 

Numeric Features : 
  
  -  TransactionDT: timedelta from a given reference datetime (not an actual timestamp).
-  TransactionAMT: transaction payment amount in USD.
-  dist1, dist2: distance
-  C1-C14: counting, such as how many addresses are found to be associated with the payment card,etc. 
The actual meaning is masked.
-  D1-D15: timedelta, such as days between previous transaction, etc.
-  Vxxx: Vesta engineered rich features, including ranking, counting, and other entity relations.
-  id_01 - id_11 : identity information 


Goal of the project:

The main goal of our project is binary classfication (i.e. two possible classes 0 or 1), we will focus on metrics commonly used for binary classifiers. Accuracy and ROC are the most commonly used evaluation metrics for binary classifiers.

### AUC(Area Under Receiver Operating Characteristic Curve)

One of the most commonly used metric to visualize the performance of a binary classifier, is the Receiver Operating Characteristic Curve(ROC) and Area Under Curve(AUC) is the best way to summarize its performance in a single number. Here, the curve refers to the ROC curve.

## Project Checklist  

Let us now go ahead with data exploration and data visualization tasks. We then proceed to data preprocessing. There are various tasks involved in this stage from feature selection to handling missing data, encoding the data and splitting the data appropriately and finallt create a model.   a.   


# Methods and Analysis   

## Data Ingestion   

In this stage, we download the data and prepare the dataset to be used in the analysis. The dataset is split into two datasets: df and validation with 90% and 10% of the original dataset respectively. We take the validation set and put it off to the side, pretending that we have never seen it before.   

```{r}
#install.packages("plyr")
#install.packages("tidyverse")
#install.packages("caret")
#install.packages("data.table")
#install.packages("lubridate")
#install.packages("utils")
#install.packages("reshape2")
#install.packages("ipred")
#install.packages("e1071")
#install.packages("pROC")
#install.packages("xgboost")
#install.packages("DMwR")
#install.packages("vip")
install.packages("randomForest")
library(plyr)
library(tidyverse)
library(caret)
library(data.table)
library(utils)
library(lubridate)
library(pROC)
library(ipred)
library(e1071)
library(xgboost)
library(reshape2)
library(xgboost)
require(DmwR)
#library(DMwR)
library(vip)
library(randomForest)
```

```{r}

# Download files 

identity <- read_csv("train_identity.csv")
transaction  <- read_csv("train_transaction.csv")
names(identity)
names(transaction)

```
We understand that there are 144233 rows and 41 columns in the identity set. 
We also understand that we have 590540 rows and 394 columns in our transaction set. 

By seeing the names of both the sets mentioned above, we understand that TransactionID column is present in both the datasets. 

If they are same, it will be easy for us to merge both the datasets and use one dataset to train the model. 

```{r}

all.equal(transaction$TransactionID,identity$TransactionID)
length(identity$TransactionID[identity$TransactionID %in% transaction$TransactionID ])

```
The output of 14423 above means that all the transaction ids present in the identity set are already present in the transaction dataset. 

So we can use left-join to merge these two datasets. 

```{r}

fulldataset<- transaction %>% 
  left_join(identity,by="TransactionID")

dim(fulldataset)

```

So, the merged dataset contains 590540rows and 434 variables. 

```{r}

# Validation set will be 10% of fulldataset data
set.seed(1, sample.kind="Rounding")
test_index <- createDataPartition(y = fulldataset$isFraud, times = 1, p = 0.1, list = FALSE)
df <- fulldataset[!row.names(fulldataset) %in% test_index,]
validation <- fulldataset[row.names(fulldataset) %in% test_index,]
validation_y <- select(validation,isFraud)
validation_x <- select(validation,-isFraud)
# Removing unwanted objects
rm(identity, transaction,fulldataset,test_index,validation)
```

## Create Helper Functions  

In this section, we will create the helper functions needed for this dataset for analysis and performance evaluation of models. The following code creates necessary helper functions for this project.  

```{r}

# This function creates one-hot encoded columns of a selected column
one_hot_encoding <- function(df,colname){
  # Get the list of possible values category can have
  possible_val <- unique(df[[colname]])
  temp_df <- cbind(sapply(possible_val, function(g) {
    as.numeric(str_detect(df[[colname]], g))
  })
  )
  return(temp_df)
}
# This function performs label encoding of the column passed
label_encoding <- function(df,colname){
  levels <- sort(unique(df[[colname]]))
  temp_df <- as.integer(factor(df[[colname]], levels = levels))
  return (temp_df)
}
# This function replaces multiple values of a column with a single value
replace_columnval <- function(col,multiple,single){
  temp <- replace(col,str_detect(col,paste(multiple,collapse = "|")),single)
  return (temp)}
```
## Exploratory Data Analysis   

### Outcome - isFraud

The response variable of this analysis is the isFraud variable. A value of 0 denotes that a transaction is fraud and 1 denotes that the transaction is legitimate. 


Let us understand the data better by seeing the distribution of fraud transactions in the given dataset. 

```{r}
# Class Distribution of Fraudulent vs Legitimate Transactions 
ggplot(data = df, aes(x = factor(isFraud), fill = factor(isFraud))) +
  geom_bar(aes(y = (..count..)/sum(..count..)),position = "dodge",alpha=0.7) + 
  geom_text(aes(y = (..count..)/sum(..count..), 
                label = paste0(..count..," ","  (",round(prop.table(..count..) * 100,2), '%',")")), 
            stat = 'count', 
            position = position_dodge(.9), 
            size=2.5,vjust = -0.3) + 
  scale_fill_manual("",values=c("#66c2a5","#fc8d62"),labels=c("Legitimate","Fraud")) +
  labs(x = "", y = '') +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "bottom", 
        axis.text=element_blank(),
        axis.ticks = element_blank())
```

From the graph, we observe that the dataset is not relatively balanced. The fraudulent transactions constitute only a small fraction(approximately 3.5%) of the dataset. 


Using this dataset as the basis for our algorithms, without deploying any techniques for balancing the dataset, can lead to errors due to overfitting. This is because the machine learning models tend to  classify all transactions as legitimate to increase the accuracy.

Hence, we need to make changes necessary and balance the data before sending it to create a model.  


When the dataset is imbalanced, it is very easy to achieve high accuracy just by classifying everything to the majority class. For example, in fraud detection datasets, majority of the transactions are legitimate and the models can achieve higher accuracy just by classifying all the transactions as legitimate. Therefore, we choose AUC for calibrating our models.

Upon analysing the data, we understand that :

## Quantitative Variables    

There are 384 quantitative predictor variables in our dataset.

The following sections discuss each of the variables in detail. 

#Transaction ID   

The transaction ID is the unique identifier of each observation in our dataset. 

# Transaction time delta   

The TransactionDT variable in our dataset gives the transaction timedelta from a given reference datetime. It does not denote an actual timestamp. The following code visualizes the total transaction amount plotted versus the timedelta for every 24 hour period from the origin.

```{r}
# Visualizing days and associated total transaction amount
df %>% 
  dplyr::mutate(Day = round(df$TransactionDT/86400)) %>%
  select(isFraud,TransactionAmt,TransactionDT,Day) %>% 
  group_by(isFraud,Day) %>% 
  dplyr::summarise(txn_amt = sum(TransactionAmt)/1000) %>% 
  ggplot(aes(Day,txn_amt,color = factor(isFraud))) +
  geom_point() +
  geom_smooth() +
  scale_color_manual("",values=c("#66c2a5","#fc8d62")) +
  facet_grid(rows = vars(isFraud), 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"))) +
  labs(x ="24-hour timedelta Relative from Origin",y="Total Transaction Amount (in Thousands)") +
  theme(legend.position = "none")
```

The graph illustrates a subtle variation in total transaction amount in a 24-hour time delta period. Moreover, the variation is similar for both legitimate and fraudulent transactions.    

# Transaction amount    

One of the most important details about a transaction is the transaction amount. The following code displays the distribution of transaction amount for legitimate and fraud transactions.  

```{r}
# Distribution of Transaction amount by fraudulent vs legitimate transactions
df %>% 
  ggplot(aes(TransactionAmt,color = factor(isFraud))) +
  geom_density() +
  facet_wrap(vars(isFraud), 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"))) +
  scale_color_manual("", 
                     values=c("#66c2a5","#fc8d62")) +
  labs(y = "Density",x = "Transaction Amount (in USD)") +
  coord_cartesian(clip = "off") +
  theme(plot.title = element_text(hjust = 0.5,face="italic"),
        legend.position = "none")


# Normalized ecdf of transaction amount of Fraudulent transactions vs Legitimate transactions
df %>% 
  ggplot(aes(x = TransactionAmt,
             y =100*..y..,
             color=factor(isFraud))) +
  stat_ecdf(geom = "step", 
            size = 0.75, 
            pad = FALSE) +
  facet_wrap(vars(isFraud), 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"))) +
  scale_x_continuous(breaks = seq(0,30000,5000)) +
  scale_color_manual("",values=c("#66c2a5","#fc8d62")) + 
  labs(x = "Transaction Amount (in USD)", y = "Cumulative Frequency") +
  coord_cartesian(clip = "off") +
  theme(legend.position = "none")

```
From the above, we undeerstand that, the distribution is a right skewed with a few transactions of higher amount, but the majority of transactions have transaction amount less than 2000 USD for legitimate transactions and 300 USD for fraudulent transactions.

## Distance    

We proceed to analyze the distance variable. The below code generates a normalized ECDF for the distance values.  

```{r}
# Visualizing dist1 and dist2 values  
df %>%
  select(dist1,dist2,isFraud) %>%
  gather("Dist","Values",-isFraud) %>% 
  ggplot(aes(Values,color = factor(isFraud))) +
  stat_ecdf(geom = "step", 
            size = 0.75, 
            pad = FALSE) +
  scale_color_manual("",values=c("#66c2a5","#fc8d62")) +
  facet_grid(isFraud~Dist, 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"))) +
  labs(x="Distance",y = "Cumulative Frequency") +
  theme(legend.position = "none")
```

We observe that, a majority of the transactions have missing distance values for dist2. However, the dist1 values are missing for 60% of the transactions. We can see that only a quarter of transactions have distance values for dist1 approximately more than 200 for fraudulent transactions.  

## C1~C14    

The variables C1 to C14 denote counting, such as how many addresses are found to be associated with the payment card,etc. The actual meaning of the variable is masked. To analyze this variable, we can examine the pairwise correlation of the variables C1~C14.

```{r}
### C1~C14 variables : counting, such as how many addresses are found to be associated with the payment card...
c_vars_df <-select(df,names(df)[str_detect(names(df),"^C[0-9]+")])
# Pairwise Correlation heatmap of C1~C14
cor(c_vars_df,
    use="pairwise.complete.obs") %>%
  melt() %>%
  dplyr::mutate(Var1=as.numeric(gsub("C([0-9]+)","\\1",Var1)),
                Var2=as.numeric(gsub("C([0-9]+)","\\1",Var2))) %>% 
  ggplot(aes(factor(Var1),factor(Var2),fill=value)) +
  geom_tile() +
  scale_fill_gradient2(low = "#d7301f",mid="white", high = "#045a8d") +
  labs(x="",y="") +
  scale_x_discrete(labels =function(x) paste0("C",x)) +
  scale_y_discrete(labels =function(x) paste0("C",x))
# Redundant features among C1~C14
c_vars_drop <- names(c_vars_df)[!str_detect(names(c_vars_df),"C([1,3,5,13]$)")]
# Remove unwanted variables
rm(c_vars_df)
```

The heatmap  illustrates the pairwise correlation between the C1~C14 variables. We see that, there is a strong correlation between many of the variables. Therefore, we consider the variables C1, C3, C5 and C13. We will exclude all other variables from our analysis.   

## D1~D15   

The variables D1 to D15 gives the timedelta, such as days between previous transaction, etc. We perform a similar approach as before and generate a pairwise correlation heatmap for the variables D1~D15.  Additionally, we eliminate the columns which have more than 85% missing data.  

```{r}
# D1 ~ D15 variables dataframe
d_vars_df <- select(df,names(df)[str_detect(names(df),"^D[0-9]+")])
# Pairwise Correlation heatmap of D1~D15
cor(d_vars_df,
    use="pairwise.complete.obs") %>%
  melt() %>%
  dplyr::mutate(Var1=as.numeric(gsub("D([0-9]+)","\\1",Var1)),
                Var2=as.numeric(gsub("D([0-9]+)","\\1",Var2))) %>% 
  ggplot(aes(factor(Var1),factor(Var2),fill=value)) +
  geom_tile() +
  scale_fill_gradient2(low = "#d7301f",mid="white", high = "#045a8d") +
  labs(x="",y="") +
  scale_x_discrete(labels =function(x) paste0("D",x)) +
  scale_y_discrete(labels =function(x) paste0("D",x))
# Remove variables with more than 90% values missing 
d_vars_missingvalues <-
  data.frame(Var = names(colMeans(is.na(df))),
             missingPercent = colMeans(is.na(df))) %>% 
  filter(grepl("^D([0-9]+)",Var)) %>% 
  filter(!round(missingPercent,2) < 0.90) %>% 
  pull(Var)
# Remove unwanted variables
rm(d_vars_df)
```

The heatmap above does not show a strong correlation between any of the variables. The variables with more than 90% of missing values can be dropped from analysis.  

#### V_ features    

 We have a total of 339 features from V1~V339.
 
 We go ahead with generating a pairwise heatmap to check for correlation between the features.  

```{r}
# Vxxx variables dataframe
vxxx_vars_df <- select(df,names(df)[str_detect(names(df),"^V")])
# Visualizing pairwise correlation of Vxxx features using heatmap
cor(vxxx_vars_df,
    use="pairwise.complete.obs") %>%
  melt() %>%
  dplyr::mutate(Var1=as.numeric(gsub("V([0-9]+)","\\1",Var1)),
                Var2=as.numeric(gsub("V([0-9]+)","\\1",Var2))) %>%
  ggplot(aes(Var1,Var2,fill=value)) +
  geom_tile() +
  scale_fill_gradient2(low = "#d7301f",mid="white", high = "#045a8d") +
  labs(x="",y="") +
  scale_x_continuous(breaks=seq(1,339,10),labels =function(x) paste0("V",x)) +
  scale_y_continuous(breaks=seq(1,339,10),labels =function(x) paste0("V",x)) +
  theme(axis.text.x = element_text(angle = 90))
# Remove variables with more than 70% values missing 
vxxx_vars_nomissingvalues <-
  data.frame(Var = names(colMeans(is.na(df))),
             missingPercent = colMeans(is.na(df))) %>% 
  filter(grepl("^V([0-9]+)",Var)) %>% 
  filter(!round(missingPercent,2) >= 0.7) %>% 
  pull(Var)
# Principal component analysis of Vxxx features
pca_Vxxx <- prcomp(na.omit(select(df,all_of(vxxx_vars_nomissingvalues))), center = FALSE, scale = FALSE)
# First 23 components explain much of the variability
#summary(pca_Vxxx)$importance[,1:16]
# Features dropped among Vxxx features
vxxx_vars_drop <- names(vxxx_vars_df)[!names(vxxx_vars_df) %in% vxxx_vars_nomissingvalues[1:16]]
# Remove unwanted variables
rm(vxxx_vars_df,pca_Vxxx,vxxx_vars_nomissingvalues)
```

Although, there appears to be a strong correlation between the V_ features, as seen above, it is quite ambiguous to separate each of the correlations. 


We can use dimensionality reduction techniques like Principal Component Analysis( PCA). The results of PCA shows that, the first 16 features account for more than 99% variance in the data. Hence, we include V1~V16 in our analysis and eliminate the others.  

## Identity Information (id_01-id_11)    

The variables id_01 to id_11 capture identity information of the transactions.

Note that the data is not availabkle for all the transactions. However,we can analyse for patterns in available data. Also , we can eliminate the features with more than 99% missing data( if there are any ).   


```{r}
# id_01 ~ id_11 :  Variables with more than 99% of the values as NA
missing_vars <- data.frame(Var = names(colMeans(is.na(df))),
                           missingPercent = colMeans(is.na(df))) %>% 
  filter(round(missingPercent,2) >= 0.99) %>%
  pull(Var)
# id_01 ~ id_38 variables dataframevariables dataframe
id_vars_df <- select(df,names(df)[str_detect(names(df),"^id_[0-9]+")])
# Remove variables with more than 99% values missing
numeric_id_vars_df <- select(df,names(select(df,names(id_vars_df)[1:11],-all_of(missing_vars))))
# Pairwise Correlation heatmap of id_01~id11
cor(numeric_id_vars_df,
    use="pairwise.complete.obs") %>%
  melt() %>%
  dplyr::mutate(Var1=as.numeric(gsub("id_([0-9]+)","\\1",Var1)),
                Var2=as.numeric(gsub("id_([0-9]+)","\\1",Var2))) %>% 
  ggplot(aes(factor(Var1),factor(Var2),fill=value)) +
  geom_tile() +
  scale_fill_gradient2(low = "#d7301f",mid="white", high = "#045a8d") +
  labs(x="",y="") +
  scale_x_discrete(labels =function(x) paste0("id_",x)) +
  scale_y_discrete(labels =function(x) paste0("id_",x))
rm(numeric_id_vars_df,id_vars_df)
```

There is no strong correlation as per figure 8. Moreover, the analysis shows that, 9 variables have more than 99% of values missing. We can eliminate these from our analysis.  


## Categorical Variables   

There are 49 categorical predictor variables in our dataset. 

We examine each one of the categorical variables one by one in this section.  

# Product Code   

The ProductCD indicates the product used in each transaction. 
Let us see if any specific product is more prone to fraudulent transactions. The following code estimates the fraction of transactions and transaction amount contributed by each of the products. The boxplot for each of the products depicts the distribution of each of the individual products.   

```{r}
# Fraudulent transactions vs legitimate transactions grouped by ProductCD
df %>% 
  group_by(ProductCD,isFraud) %>%
  dplyr::summarise(num_txn = n()) %>%
  data.frame() %>% 
  dplyr::mutate(prop = num_txn/sum(num_txn)) %>% 
  ggplot(aes(x = ProductCD,y = prop, fill = factor(isFraud))) + 
  geom_bar(stat="identity",
           position=position_dodge(),
           alpha=0.7,
           color="white") +
  geom_text(aes(x = ProductCD, y = prop,
                label = paste0(round(prop * 100,2), '%')),
            position=position_dodge(0.9), 
            size=2.5,vjust = -1) +
  labs(x = "Product Code",y = "% of Number of Transactions") +
  scale_y_continuous(labels = scales::percent,limits = c(0,1)) +
  scale_fill_manual("",values=c("#66c2a5","#fc8d62"),labels=c("Legitimate","Fraud")) +
  theme(legend.position = "bottom")
# Total Transaction amount of Fraudulent transactions vs legitimate transactions grouped by ProductCD
df %>% 
  group_by(isFraud,ProductCD) %>%
  dplyr::summarise(Txn_amount = sum(TransactionAmt)/1000) %>% 
  ggplot(aes(x = ProductCD,y = Txn_amount, fill = factor(isFraud))) + 
  geom_bar(stat="identity",
           position=position_dodge(),
           alpha=0.7,
           color="white") +
  geom_text(aes(x = ProductCD, y = Txn_amount,
                label = paste0(round(Txn_amount,2),"$")),
            position=position_dodge(0.9), 
            size = 2,vjust=-0.1) +
  facet_wrap(ProductCD ~.,scales = "free") +
  scale_fill_manual("",values=c("#66c2a5","#fc8d62"),labels =c("Legitimate","Fraud")) +
  labs(y="Total Transaction Amount (in Thousands)",x = "") +
  theme(legend.position = "bottom")
# Distribution of Transaction amount of Fraudulent transactions vs legitimate transactions grouped by ProductCD
df %>% 
  ggplot(aes(x = ProductCD,y = TransactionAmt, fill = factor(isFraud))) + 
  geom_boxplot(outlier.colour="red",
               outlier.shape=1,
               alpha=0.7) +
  labs(y="Transaction Amount") +
  ylim(0,1500) +
  coord_flip() +
  scale_fill_manual("",
                    values=c("#66c2a5","#fc8d62"),
                    labels = c("Legitimate", "Fraud")) +
  theme(legend.position = "bottom")
```

We can analyse the number of transactions in a particular product through this and we retain this ProductCD feature .

#### Payment Card Information    

The payment card information includes 6 variables card1~card6. The information in these variable contain details such as card type, card category, issue bank, country, etc.  

```{r}

# Visualizing  variables Card 1, Card2 and Card3 and Card5 and associated number of transactions
df %>% 
  select(isFraud,card1,card2,card3,card5) %>%
  gather("Card","Values",-isFraud) %>% 
  group_by(Card,isFraud,Values) %>%
  dplyr::summarise(num_txn = n()) %>%
  ggplot(aes(x = Values,y = num_txn,color=factor(Card))) +
  geom_point(alpha=0.7) +
  scale_color_manual("",values=c("#66c2a5","#fc8d62","#8da0cb","#ffd92f","#bc80bd")) +
  facet_wrap(Card ~ isFraud, 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud")),
             ncol=2) +
  labs(x ="card1 values",y="Number of Transactions") +
  theme(legend.position = "none")
# Visualizing  card1, card2, card3 and card5 values and associated transaction amount
df %>% 
  select(isFraud,TransactionAmt,card1,card2,card3,card5) %>%
  gather("Card","Values",-isFraud,-TransactionAmt) %>% 
  group_by(Card,isFraud,Values) %>%
  dplyr::summarise(txn_amt = sum(TransactionAmt)/1000) %>%
  ggplot(aes(x = Values,y = txn_amt,color=factor(Card))) +
  geom_point(alpha=0.7) +
  scale_color_manual("",values=c("#66c2a5","#fc8d62","#8da0cb","#ffd92f","#bc80bd")) +
  facet_wrap(Card ~ isFraud, 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud")),
             ncol=2) +
  labs(x ="card1 values",y="Number of Transactions") +
  theme(legend.position = "none")
```

card4-6:


```{r}
# Visualizing variables card4 and card6 
# Number of transactions for each Card type and network, grouped by transaction type(Fraudulent/legitimate)
df %>%
  dplyr::mutate(card4 = replace_columnval(card4,
                                          unique(df$card4)[!str_detect(unique(df$card4),
                                                                       "visa|master|discover|amer")],"Others"),
                card6 = replace_columnval(card6,unique(df$card6)[!str_detect(unique(df$card6),
                                                                             "debit|credit")],"Others")) %>%
  group_by(isFraud,card4,card6) %>%
  dplyr::summarise(num_txn = n()) %>%
  ggplot(aes(y = num_txn,x = card6, fill=factor(card6))) +
  geom_bar(stat="identity",alpha=0.8,position=position_dodge()) +
  facet_wrap(isFraud ~ card4 , 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"),
                                 card4 = c("Other Networks"= "Other \n Networks","american express"="Amex","discover"="Discover","mastercard"="MasterCard","visa"="Visa")),
             ncol = 5) +
  scale_fill_manual("",values=c("#66c2a5","#fc8d62","#8da0cb","#ffd92f","#bc80bd")) +
  labs(y="Number of Transactions",x = "Card Type") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))
# Total transaction amount for each Card type and network, grouped by transaction type(Fraudulent/legitimate)
df %>% 
  dplyr::mutate(card4 = replace_columnval(card4,
                                          unique(df$card4)[!str_detect(unique(df$card4),
                                                                       "visa|master|discover|amer")],"Others"),
                card6 = replace_columnval(card6,unique(df$card6)[!str_detect(unique(df$card6),
                                                                             "debit|credit")],"Others")) %>% 
  group_by(isFraud,card4,card6) %>%
  dplyr::summarise(Txn_amount = sum(TransactionAmt)/1000) %>%
  ggplot(aes(y = Txn_amount,x = card6, fill=factor(card6))) +
  geom_bar(stat="identity",alpha=0.8) +#position=position_dodge()
  facet_wrap(isFraud ~ card4 , 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"),
                                 card4 = c("Other Networks"= "Other Networks","american express"="Amex","discover"="Discover","mastercard"="MasterCard","visa"="Visa")),
             ncol = 5) +
  labs(y="Transaction Amount (in Thousands)",x = "Card Type") +
  scale_fill_manual("",values=c("#66c2a5","#fc8d62","#8da0cb","#ffd92f","#bc80bd")) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))
```

This information is essential to analyse the genuinity of the transactions and hence we retain all these features. 

## Match   
M1~M9. 

These features denote match, such as names on card and address, etc. 
The number of transactions and total amount associated with each kind of value can be viewed in the plots below.  

Except for M4 variables, all other match variables are of boolean type. We will visualize both of them separately to look for patterns to identify transactions as fraud.  

```{r}
# Visualizing M1~ M9 values and associated number of transactions
df %>% 
  select(isFraud,M1:M3,M5:M9) %>%
  gather("Var","values",-isFraud) %>%
  group_by(Var,isFraud,values) %>%
  dplyr::summarise(num_txn = n()) %>%
  ggplot(aes(x = Var,y = num_txn,fill=factor(Var))) +
  geom_bar(stat="identity",
           alpha=0.65,
           position = position_dodge()) +
  scale_fill_brewer(type="qual",palette=3) +
  facet_wrap(values ~ isFraud, 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud")),
             ncol=2) +
  labs(x ="",y="Number of Transactions") +
  theme(legend.position = "bottom",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
# Visualizing M1~M9 values and associated total transaction amount
df %>% 
  select(TransactionAmt,isFraud,M1:M3,M5:M9) %>%
  gather("Var","values",-TransactionAmt,-isFraud) %>%
  group_by(Var,isFraud,values) %>%
  dplyr::summarise(txn_amt = sum(TransactionAmt)/1000) %>%
  ggplot(aes(Var,y=txn_amt,fill=factor(Var))) +
  scale_fill_brewer(type="qual",palette=3) +
  geom_bar(stat="identity",
           alpha=0.65,
           position = position_dodge()) +
  facet_wrap(values ~ isFraud, 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud")),
             ncol=2) +
  labs(x="",y="Total Transaction Amount (in Thousands)") +
  theme(legend.position = "bottom")
### M4 (Character variable) 
# Visualizing M4 values and associated number of transactions 
df %>% 
  select(isFraud,M4) %>%
  dplyr::mutate(M4 = replace(M4,is.na(M4),"(Blank)")) %>% 
  group_by(isFraud,M4) %>%
  dplyr::summarise(num_txn = n()) %>%
  ggplot(aes(x = M4,y = num_txn,color = factor(isFraud),fill=factor(M4))) +
  geom_bar(stat="identity",
           position = position_dodge()) +
  scale_fill_brewer(type="qual",palette=1) +
  scale_color_manual("",values=c("#66c2a5","#fc8d62")) +
  facet_wrap(vars(isFraud), 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud")),
             ncol=2) +
  labs(x ="M4 values",y="Number of Transactions") +
  theme(legend.position = "none")
# Visualizing M4 values and associated total transaction amount
df %>% 
  select(TransactionAmt,isFraud,M4) %>%
  dplyr::mutate(M4 = replace(M4,is.na(M4),"(Blank)")) %>% 
  group_by(isFraud,M4) %>%
  dplyr::summarise(txn_amt = sum(TransactionAmt)/1000) %>%
  ggplot(aes(x = M4,y = txn_amt,color = factor(isFraud),fill=factor(M4))) +
  geom_bar(stat="identity",
           position = position_dodge()) +
  scale_fill_brewer(type="qual",palette=1) +
  scale_color_manual("",values=c("#66c2a5","#fc8d62")) +
  facet_wrap(vars(isFraud), 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud")),
             ncol=2) +
  labs(x ="M4 values",y="Total Transaction Amount (in Thousands)") +
  theme(legend.position = "none")
```
There is variance accounted by each of these features. We retain all these features for modeling.  

#### Address    

The address details are given by addr1 and addr2 columns. We examine addr1 and addr2 separately to look for discernible patterns.   

```{r}
# Visualizing addr1 values and associated number of transactions
df %>%
  group_by(isFraud,addr1) %>%
  dplyr::summarise(num_txn = n()) %>%
  ggplot(aes(addr1,num_txn,color=factor(isFraud))) +
  geom_point() +
  scale_color_manual("",values=c("#66c2a5","#fc8d62")) +
  facet_grid(rows = vars(isFraud), 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"))) +
  labs(x ="addr1 values",y="Number of Transactions") +
  theme(legend.position = "none")
# Visualizing addr1 values and associated total transaction amount
df %>%
  group_by(isFraud,TransactionAmt,addr1) %>%
  dplyr::summarise(txn_amt = sum(TransactionAmt)) %>%
  ggplot(aes(addr1,txn_amt,color=factor(isFraud))) +
  geom_point() +
  scale_color_manual("",values=c("#66c2a5","#fc8d62")) +
  facet_grid(rows = vars(isFraud), 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"))) +
  labs(x ="addr1 values",y="Total Transaction Amount (in Thousands)") +
  theme(legend.position = "none")
```
```{r address2, echo = TRUE, fig.cap="Visualizing addr2",fig.subcap=c('Number of transactions' ,'Total transaction amount'), fig.ncol=2}
# Visualizing addr2 values and associated number of transactions
df %>%
  group_by(isFraud,addr2) %>%
  dplyr::summarise(num_txn = n()) %>%
  ggplot(aes(addr2,num_txn,color=factor(isFraud))) +
  geom_point() +
  scale_color_manual("",values=c("#66c2a5","#fc8d62")) +
  geom_vline(xintercept=87,linetype = "longdash",color="grey") +
  geom_text(aes(87,y=-500,label=87,hjust=1.5),
            size=2.5,family="Courier", fontface="italic",color="darkgrey") +
  facet_grid(rows = vars(isFraud), 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"))) +
  labs(x ="addr2 values",y="Number of Transactions") +
  theme(legend.position = "none")
# Visualizing addr2 values and associated total transaction amount
df %>%
  group_by(isFraud,TransactionAmt,addr2) %>%
  dplyr::summarise(txn_amt = sum(TransactionAmt)) %>%
  ggplot(aes(addr2,txn_amt,color=factor(isFraud))) +
  geom_point() +
  scale_color_manual("",values=c("#66c2a5","#fc8d62")) +
  geom_vline(xintercept=87,linetype = "longdash",color="grey") +
  geom_text(aes(87,y=-500,label=87,hjust=1.5),
            size=2.5,family="Courier", fontface="italic",color="darkgrey") +
  facet_grid(rows = vars(isFraud), 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"))) +
  labs(x ="addr2 values",y="Total Transaction Amount (in Thousands)") +
  theme(legend.position = "none")
```

#### Purchaser email domain   

The purchaser email domain is given by P_emaildomain. We can merge the domains appropriately to ease our analysis with domains. As there are some empty fields in these column, we will merge all microsoft, yahoo, icloud, aol, gmail and .net(Networking) domains separately and categorize all other domains as "Others".   

```{r}
# Visualizing purchaser email domain of approximately 99% of the transactions.
df %>%
  dplyr::mutate(P_emaildomain = replace_columnval(P_emaildomain,c("live","outlook","hotmail","msn"),"Microsoft"),
                # Replaces 8 Yahoo domains 
                P_emaildomain = replace_columnval(P_emaildomain,c("yahoo","ymail"),"Yahoo"),
                # Replaces 3 icloud domains 
                P_emaildomain = replace_columnval(P_emaildomain,c("me.com","mac.com","icloud.com"),"iCloud"),
                # Replaces 2 AOL domains 
                P_emaildomain = replace_columnval(P_emaildomain,c("aol","aim"),"AOL"),
                P_emaildomain = replace_columnval(P_emaildomain,c("gmail.com","gmail"),"Gmail"),
                P_emaildomain = replace_columnval(P_emaildomain,c(".net"),"NetworkingDomains"),
                P_emaildomain = replace_columnval(P_emaildomain,unique(df$P_emaildomain)[!str_detect(unique(df$P_emaildomain),
                                                                                                     "Gmail|Yahoo|Networking|Microsoft|AOL|iCloud")]
                                                  ,"Others")) %>%
  dplyr::mutate(P_emaildomain = replace(P_emaildomain,is.na(P_emaildomain),"Unknown")) %>%
  group_by(isFraud,P_emaildomain) %>%
  dplyr::summarise(num_txn = n()) %>%
  ggplot(aes(reorder(P_emaildomain,num_txn),num_txn,fill=factor(isFraud))) +
  geom_bar(stat="identity",position=position_dodge()) +
  scale_fill_manual("",values=c("#66c2a5","#fc8d62")) +
  facet_wrap(vars(isFraud), 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"))) +
  coord_flip() +
  labs(x ="Purchaser email domain",y="Number of transactions") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))
```
From the above, we can understand hat the number of genuine transactions for yahoo exceeds all others except gmail. But, the number of fraud transactions from microsoft exceeds that of yahoo. There are some patterns which need more analysis and hence we retain the variable in our model. 


##Recepient email domain   

The recepient email domain is given by R_emaildomain. We use the same technique that we used for purchaser email domain above. 

```{r}
# Visualizing recipient email domain of approximately 99% of the transactions.
df %>%
  dplyr::mutate(R_emaildomain = replace_columnval(R_emaildomain,c("live","outlook","hotmail","msn"),"Microsoft"),
                R_emaildomain = replace_columnval(R_emaildomain,c("yahoo","ymail"),"Yahoo"),
                R_emaildomain = replace_columnval(R_emaildomain,c("me.com","mac.com","icloud.com"),"iCloud"),
                R_emaildomain = replace_columnval(R_emaildomain,c("aol","aim"),"AOL"),
                R_emaildomain = replace_columnval(R_emaildomain,c("gmail.com","gmail"),"Gmail"),
                R_emaildomain = replace_columnval(R_emaildomain,c(".net"),"NetworkingDomains"),
                R_emaildomain = 
                  replace_columnval(R_emaildomain,
                                    unique(df$R_emaildomain)[!str_detect(unique(df$R_emaildomain),
                                                                         "Gmail|Yahoo|Networking|Microsoft|AOL|iCloud")],"Others")) %>%
  dplyr::mutate(R_emaildomain = replace(R_emaildomain,is.na(R_emaildomain),"Unknown")) %>%
  group_by(isFraud,R_emaildomain) %>%
  dplyr::summarise(num_txn = n()) %>%
  ggplot(aes(reorder(R_emaildomain,num_txn),num_txn,fill=factor(isFraud))) +
  geom_bar(stat="identity",position=position_dodge()) +
  scale_fill_manual("",values=c("#66c2a5","#fc8d62")) +
  facet_wrap(vars(isFraud),
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"))) +
  coord_flip() +
  labs(x ="Recepient email domain",y="Number of transactions") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))
```
We see similar patters with recipient email domain as well. Even though number of genuine transactions are higher in aol when compared to icloud, the latter has higher fraud transactions.  

## Device Type   

The variable DeviceType gives the type of device used for the online transaction. We shall examine the number of transactions in each category.  

```{r}
# Device Type : type of device used for transaction (mobile/desktop/Unknown)
df %>%
  dplyr::mutate(DeviceType = replace(DeviceType,is.na(DeviceType),"Unknown")) %>%
  group_by(isFraud,DeviceType) %>% 
  dplyr::summarise(num_txn = n()) %>%
  ggplot(aes(y = num_txn,x = DeviceType, fill=factor(isFraud))) +
  geom_bar(stat="identity",alpha=0.8,position=position_dodge()) +
  facet_wrap(vars(isFraud), 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"))) +
  scale_fill_manual("",values=c("#66c2a5","#fc8d62")) +
  labs(y="Number of Transactions",x = "Device Type") +
  theme(legend.position = "none")
```
Although the number of genuine transactions using desktop exceeds that of mobile, the number of fraud transactions using mobile is slightly higher than that of desktop. We shall retain the variable for inclusion in our model.  

## Device Info   

The device information of the device on which the transaction was made can is available in the DeviceInfo column. We analyse the device information of devices that have more than 200 transactions.  
```{r}
### Device Info 
df %>%
  select(DeviceType,DeviceInfo,isFraud) %>%
  dplyr::mutate(DeviceInfo = replace_columnval(DeviceInfo,c("SM","SAMSUNG","samsung"),"Samsung"),
                DeviceInfo = replace_columnval(DeviceInfo,c("HTC"),"HTC"),
                DeviceInfo = replace_columnval(DeviceInfo,c("Moto","Motorola","MOTOROLA","XT1635"),"Motorola"),
                DeviceInfo = replace_columnval(DeviceInfo,c("HUAWEI","Huawei","hi6210"),"Huawei"),
                DeviceInfo = replace_columnval(DeviceInfo,c("LG"),"LG"),
                DeviceInfo = replace(DeviceInfo,is.na(DeviceInfo),"Unknown")) %>%
  dplyr::mutate(DeviceType = replace(DeviceType,is.na(DeviceType),"Unknown")) %>% 
  group_by(DeviceType,DeviceInfo,isFraud) %>%
  dplyr::summarise(num=n()) %>%
  arrange(-num) %>% 
  filter(num>200) %>%  
  ggplot(aes(y = num,x = DeviceInfo, fill=factor(isFraud))) +
  geom_bar(stat="identity",alpha=0.8,position=position_dodge()) +
  facet_wrap(isFraud~DeviceType, 
             scales = "free",
             labeller = labeller(isFraud = c("0"= "Legitimate","1" = "Fraud"),
                                 DeviceType = c("desktop"="Desktop","mobile"="Mobile","Unknown"="Unknown"))) +
  scale_fill_manual("",values=c("#66c2a5","#fc8d62","#8da0cb")) +
  labs(y="Number of Transactions",x = "Device Information") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))
```
There are visible patterns in the visualization above and hence we use this variable in our model.  

#### Identity Information (id_12~id_38)   

The identity information available in id12~id38 contain information regarding the device where transaction was made. For example : id_30 gives the OS version, id_31 gives the browser version, id_33 gives the resolution of the device to name a few. We retain all these variables, since they give further details of the transaction.

## Data Preprocessing   

This stage involves various data preprocessing tasks like handling missing data, encoding categorical data as numerical and imputing missing values.


Feature Selection and Imputing Missing data :

We already know that we had 434 features/variables in the original dataset. 
During our data analysis, taking many factor into consideration like missing values, correlation etc., we thought of eliminating few features. 

The following is the code to eliminate before using the dataset for creation of a model. 
```{r}
eliminated_features <-c(c_vars_drop,d_vars_missingvalues,vxxx_vars_drop,missing_vars)
rm(c_vars_drop,d_vars_missingvalues,vxxx_vars_drop,missing_vars)

```


Label Encoding and One hot encoding:

We use this to encode the dataset and impute missing values where required. 


```{r}
feature_engg_fraud_det <- function(temp){
  # Eliminated 350 variables after EDA; Remaining 84 variables
  numeric_df <- select(temp,-all_of(eliminated_features)) %>% select_if(is.numeric)  #  Numeric vectors
  char_df <- select(temp,-all_of(eliminated_features)) %>% select_if(is.character)   #  Character vectors
  logical_df <- select(temp,-all_of(eliminated_features)) %>% select_if(is.logical) # 12 Logical vectors
  
  # Merging email domains, card values and imputing missing categorical values
  char_df <- char_df %>%
    dplyr::mutate(card6 = replace_columnval(card6,
                                            unique(char_df$card6)[!str_detect(unique(char_df$card6),
                                                                              "debit|credit")],"Others")) %>%
    dplyr::mutate(card4 = replace_columnval(card4,
                                            unique(char_df$card4)[!str_detect(unique(char_df$card4),
                                                                              "visa|master|discover|amer")],
                                            "Others")) %>%
    dplyr::mutate(P_emaildomain = replace_columnval(P_emaildomain,
                                                    c("live","outlook","hotmail","msn"),
                                                    "Microsoft"),
                  # Replaces Yahoo domains 
                  P_emaildomain = replace_columnval(P_emaildomain,
                                                    c("yahoo","ymail"),
                                                    "Yahoo"),
                  # Replaces icloud domains 
                  P_emaildomain = replace_columnval(P_emaildomain,c("me.com","mac.com","icloud.com"),"iCloud"),
                  # Replaces AOL domains 
                  P_emaildomain = replace_columnval(P_emaildomain,
                                                    c("aol","aim"),
                                                    "AOL"),
                  P_emaildomain = replace_columnval(P_emaildomain,
                                                    c("gmail.com","gmail"),
                                                    "Gmail"),
                  P_emaildomain = replace_columnval(P_emaildomain,
                                                    c(".net"),
                                                    "NetworkingDomains"),
                  P_emaildomain = replace_columnval(P_emaildomain,
                                                    unique(char_df$P_emaildomain)[!str_detect(unique(char_df$P_emaildomain),
                                                                                              "Gmail|Yahoo|Networking|Microsoft|AOL|iCloud")]
                                                    ,"Others")) %>%
    dplyr::mutate(R_emaildomain = replace_columnval(R_emaildomain,
                                                    c("live","outlook","hotmail","msn"),
                                                    "Microsoft"),
                  R_emaildomain = replace_columnval(R_emaildomain,
                                                    c("yahoo","ymail"),
                                                    "Yahoo"),
                  R_emaildomain = replace_columnval(R_emaildomain,
                                                    c("me.com","mac.com","icloud.com"),
                                                    "iCloud"),
                  R_emaildomain = replace_columnval(R_emaildomain,
                                                    c("aol","aim"),
                                                    "AOL"),
                  R_emaildomain = replace_columnval(R_emaildomain,
                                                    c("gmail.com","gmail"),
                                                    "Gmail"),
                  R_emaildomain = replace_columnval(R_emaildomain,
                                                    c(".net"),
                                                    "NetworkingDomains"),
                  R_emaildomain = replace_columnval(R_emaildomain,
                                                    unique(char_df$R_emaildomain)[!str_detect(unique(char_df$R_emaildomain),
                                                                                              "Gmail|Yahoo|Networking|Microsoft|AOL|iCloud")]
                                                    ,"Others")) %>%
    dplyr::mutate(DeviceType =  replace(DeviceType,is.na(DeviceType),"Unknown"),
                  DeviceInfo =  replace(DeviceInfo,is.na(DeviceInfo),"Unknown"),
                  card4 = replace(card4,is.na(card4),"Unknown"),
                  card6 = replace(card6,is.na(card6),"Unknown"),
                  P_emaildomain = replace(P_emaildomain,is.na(P_emaildomain),"Unknown"),
                  R_emaildomain = replace(R_emaildomain,is.na(R_emaildomain),"Unknown"),
                  M4 =  replace(M4,is.na(M4),"Unknown"),
                  id_12 = replace(id_12,is.na(id_12),"Unknown"),
                  id_15 = replace(id_15,is.na(id_15),"Unknown"),
                  id_16 = replace(id_16,is.na(id_16),"Unknown"),
                  id_28 = replace(id_28,is.na(id_28),"Unknown"),
                  id_29 = replace(id_29,is.na(id_29),"Unknown"),
                  id_30 = replace(id_30,is.na(id_30),"Unknown"),
                  id_31 = replace(id_31,is.na(id_31),"Unknown"),
                  id_34 = replace(id_34,is.na(id_34),"Unknown")) %>%
    # Dropping resolution column
    select(-id_33)
  # Binding columns of the dataset :  106 variables
  temp_df <-
    # 5 character vectors encoded as 28 one-hot encoded variables
    cbind(as.data.frame(sapply(names(char_df)[str_detect(names(char_df),
                                                         "emaildomain|card|Product")],
                               function(x) one_hot_encoding(char_df,x))),
          # 11 character variables encoded with label encoding to 11 variables
          as.data.frame(sapply(names(char_df)[!str_detect(names(char_df),
                                                          "emaildomain|card|Product")],
                               function(x) label_encoding(char_df,x))),
          # 12 logical vectors encoded as 12 numeric encoding of logical variables
          as.data.frame(sapply(names(logical_df), 
                               function(x) as.numeric(logical_df[[x]]))),
          # 55 numeric variables 
          numeric_df)
  
  # Imputing numerical NA as -999
  temp_df <- temp_df %>% dplyr::mutate_if(is.numeric, list(~replace_na(., -999)))
  
  return (temp_df)
}
```

We then create a 90/10 train-test split of the df dataset for our model evaluation purposes. 
The model is created and trained in the train set and tested in the test set. 

```{r}
# Create train set, test set (Imbalanced)
df <- feature_engg_fraud_det(df) 
set.seed(1234, sample.kind="Rounding") 
# Create train set and test set
test_index <- createDataPartition(y = df$isFraud, times = 1 ,p=0.1, list = FALSE)
imbal_train_set <- df[!row.names(df) %in% test_index,]
imbal_test_set <- df[row.names(df) %in% test_index,]
# Remove unwanted variables
rm(test_index)
```

## Dealing with Imbalanced Classes

We previously discussed the effects of imbalanced data on our machine learning algorithms. Algorithms tend to overfit the data and classify all the transactions as legitimate and fail to discern patterns in the fraudulent transactions. However, there are methods to mitigate this and balance the data before feeding it to the models. 
We can achieve balanced classes by downsampling the majority class or oversampling the minority class.
Although, with random undersampling method there is a loss in information of the non-fraud transactions, it is optimal method with regards to computational efficiency. Oversampling the minority class can be achieved using Synthetic Minority Oversampling Technique, or SMOTE. With our current dataset, deploying SMOTE to balance the dataset needs excessive computational resources. 
Therefore, we proceed with random downsampling. In this method, the random samples are chosen from majority class and all minority class transactions in the dataset is retained.



Training model woth the final dataset:

#RANDOM FOREST:


Random forests build an ensemble of deep independent trees to improve the predictive performance. Random forests are an improvement over bagged decision trees. The fundamental principle behind random forests are same as that of decision trees. They help to minimize the tree correlation by randomly picking the bootstrapped samples. The algorithm performs split-variable randomization where each time a split is to be performed and the search for the split variable is controlled by the mtry parameter. The default value of mtry is $\sqrt{p}$ for classification model.   

#EXtreme Gradient Boosting(XGBOOST)   

Gradient Boosting Techniques are extremely popular machine learning algorithms and have proven successful across various domains. Gradient Boosting Machines(GBMs) build an ensemble of shallow trees in sequence, with each tree learning from the previous and are highly efficient, flexible and portable.

# Results    

This section presents the code and results of the models discussed in the previous section. We will measure AUC for each of the models discussed above. The trainControl function is used to set the parameters to control how models are created. The models we chose in the previous section will be used along with these parameters. We will use repeated cross validation technique with 5-fold cross validation and the sampling technique used is the random down sampling technique. The classProbs is set to TRUE to enable the models to compute class probabilities, so that, AUC can be estimated.   


```{r ctrl_params,echo=TRUE}
# Computational control parameters for train function
fitControl = trainControl(method = "repeatedcv", # Repeated Cross validation 
                          number = 5,  #  controls the number of folds in K-fold cross-validation
                          #repeats = 5, # Repeats K-fold cross-validation
                          verboseIter = FALSE, # A logical for printing a training log.
                          returnData = FALSE, #  A logical for saving the data into a slot called trainingData
                          allowParallel = TRUE, # Use a parallel processing if available
                          sampling = "down", # Balancing imbalanced dataset using random downsampling
                          classProbs = TRUE, #  class probabilities be computed for classification models
                          summaryFunction = twoClassSummary) # a function to computed alternate performance summaries.
```


### Random Forest   

In this section, we implement random forest algorithm for our fraud detection data. The mtry hyperparameter that controls split-variable randomization feature of random forests has been set to square root of the number of variables in our model.  

```{r randomforest,echo=TRUE}
# Model 1: Random Forest
# mtry (The number of features to consider at any given split) for Classification models 
mtry = floor(sqrt(ncol(imbal_train_set)))
# Creating hyperparameter grid search for Random Forest
rfgrid <- expand.grid(.mtry=mtry)
# Setting seed for reproducibility 
set.seed(112, sample.kind="Rounding") 
# Train the model to find the best tune for Random Forest
fit.randomforest = train(
  make.names(isFraud)~.,
  data = imbal_train_set,  
  trControl = fitControl,
  tuneGrid = rfgrid,
  method = "rf",
  tree_method = "hist",
  metric ="ROC",
  verbose = TRUE)
# Predict with imbalanced test set
pred.rf <- predict(fit.randomforest,
                   imbal_test_set %>% select(-isFraud),
                   type = "prob")
```


```{r}
# Generating ROC for the Random Forest model
rf_down <- roc(imbal_test_set$isFraud,
               factor(pred.rf$X1,ordered=TRUE),
               plot=TRUE,
               print.auc=TRUE,
               col = "#6a51a3",  
               legacy.axes=TRUE,
               levels = c("0", "1"))
rf_down
```
 AUC(= 0.941) with the random forest algorithm is noted.

Variable importance for the random forest fit model:

```{r}
vip::vip(fit.randomforest, num_features = 40, geom = "point")
```

### eXtreme Gradient Boosting(XGBOOST)      

The following code accomplishes the XGBOOST with best tune parameters, obtained after a grid search on a hyperparameter grid of values. Boosting iterations has been restricted to 500 for optimal computational efficiency.

```{r}
# XGBOOST (eXtreme Gradient Boosting)
# Creating hyperparameter grid search for XGBOOST Hyperparameters with faster learning rate of 0.1
xgbGrid = expand.grid(nrounds = 500,   # Boosting Iterations : Number of trees or rounds
                      max_depth = 15,  # c(5, 10, 15), # Max Tree Depth : Controls the depth of the individual trees
                      eta = 0.1,       # Learning rate/ Shrinkage
                      gamma = 1,       # c(1, 2, 3), # Minimum Loss Reduction : Pseudo-regularization hyperparameter known as a Lagrangian multiplier and controls the complexity of a given tree.
                      colsample_bytree = 1, # c(0.4, 0.7, 1.0), # Subsample Ratio of Columns
                      min_child_weight = 0.5, # c(0.5, 1, 1.5), # Minimum Sum of Instance Weight
                      subsample = 0.7 ) # Subsample Percentage

# Setting seed for reproducibility
set.seed(212, sample.kind="Rounding") 
# Train the model to find the best tune for XGBOOST
fit.xgboost = train(
  make.names(isFraud)~.,
  data = imbal_train_set,  
  trControl = fitControl,
  tuneGrid = xgbGrid,
  method = "xgbTree",
  tree_method = "hist",
  metric ="ROC",
  verbose = TRUE)
# Predict with imbalanced test set
pred.xgboost <- predict(fit.xgboost,
                        imbal_test_set %>% select(-isFraud),
                        type = "prob")
```

```{r}
# Generating ROC for the XGBOOST model
xgb_down <- roc(imbal_test_set$isFraud,
                factor(pred.xgboost$X1,ordered=TRUE),
                plot=TRUE,
                print.auc=TRUE,
                col = "#993404", 
                legacy.axes=TRUE,
                levels = c("0", "1"))
xgb_down
```


We see improvement in AUC with the XGBOOST algorithm. The AUC achieved by the XGBOOST model corresponds to `r round(xgb_down$auc,5)`. We then proceed to construct a variable importance plot (VIP) of the top 40 features in the XGBOOST model.  

```{r}
vip::vip(fit.xgboost, num_features = 40, geom = "point")
```

## Final Validation   

Once we are confident about our final model, we measure AUC on the validation set, to estimate the generalization performance. 

In the case of fraud detection dataset, the finalized model is XGBOOST model. Now, we train the complete model with the df dataset and estimate the AUC on validation set.  


```{r}
### AUC of finalized model(XGBOOST) with the validation set 
# Set random seed 
#set.seed(1234, sample.kind = "Rounding")
# Training the full model with the best tuning hyperparameters
#fxgb = train(
 # make.names(isFraud)~.,
 # data = df,
  #trControl = fitControl,
  #tuneGrid = xgbGrid,
  #method = "xgbTree",
  #tree_method = "hist",
  #metric ="ROC",
  #verbose = TRUE)
# Predictions on the validation set
#pred.xgboost <- predict(fxgb,
                        #feature_engg_fraud_det(validation_x),
                        #type = "prob")
# Generating ROC Curve for XGBOOST on validation dataset
#rocxgb_down_validation <- roc(validation_y$isFraud,
                              #factor(pred.xgboost$X1,ordered=TRUE),
                              #plot=TRUE,
                              #print.auc=TRUE,
                              #col = "#006d2c", # green
                              #legacy.axes=TRUE,
                              #main="AUC with XGBOOST(Validation Set)","""
                             # levels = c("0", "1"))
#rocxgb_down_validation



#auc_results <- 
  #rbind(auc_results,tibble(method = "XGBOOST (Validation Set)", 
                           #AUC =rocxgb_down_validation$auc))

#auc_results %>% knitr::kable()
```

The AUC achieved by the XGBOOST on validation set corresponds to `r round(rocxgb_down_validation$auc,5)`.
We then proceed to construct a variable importance plot (VIP) of the top 40 features in the XGBOOST model on validation set.

```{r}
#vip::vip(fxgb, num_features = 40, geom = "point")
```


# Conclusion    

We have used Random forest and XGBoost algorithms to find our best model.Regardless of the the computational intensity of the algorithms discussed above, it is vital to note that gradient boosting machines are one of the most powerful ensemble algorithms offering excellent accuracy.


```{r}
# Removing unwanted objects
#rm(df,validation_x,validation_y,auc_results,rocxgb_down_validation,xgb_down)
```

